Rapid-response Bayesian sky localization for electromagnetic follow-up of gravitational-wave candidates

### 2013-02-28 ms.tex
Fix build error
Use boldface rather than overarrows for vectors

The quantity of interest for any type of parameter estimation is the so-called {\it posterior} distribution, $p(\vec{\theta}|d)$, that describes the probability of obtaining the parameters, $\vec{\theta}$, given the data, $d$.  It is given by Bayes' theorem
\begin{equation}\label{bayes}
p(\vec{\theta}|d) = \frac{p(\vec{\theta})p(d|\vec{\theta})}{p(d)},
\end{equation}
where $p(\vec{\theta})$ is the {\it prior} distribution describing previous knowledge of the paramters, $\vec{\theta}$, $p(d|\vec{\theta})$ is the {\it likelihood} of obtaining the data, $d$, for the parameters, $\vec{\theta}$, and $p(d)$ is known as the {\it evidence}.  Procedures for computing the posterior typically rely on Markov chain Monte Carlo methods or nested sampling (something else?).  In this manuscript we present an alternative approach, taylored specifically to providing low-latency estimates of the sky location of and distance to the source.

In the full signal model the paramters, $\vec{\theta}$, are given by a 15-dimensional vector of quantities.  Of those, the quantities of interest are the right ascenion and declination, $(\alpha,\delta)$, and luminosity distance, $D_L$.  The remaining quantites, or {\it nuisance parameters} for the full signal model are:
\begin{itemize}
\item The chirp mass.  If $m_1$ and $m_2$ are the masses of the individual components then $\mathcal{M}=(m_1m_2)^{3/5}/(m_1+m_2)^{1/5}$.
\item The symmetric mass ratio, $\eta=(m_1m_2)/(m_1+m_2)^2$.
\item The 3-dimensional spin vectors of each components, $\vec{S}_1,\vec{S}_2$.
\item An arrival time, $\tau_\oplus$, taken here to be the time of arrival at the geocenter.
\item The inclination angle, $\iota$.
\item The polarization angle, $\psi$.
\item The coalescence phase, $\phi_c$.
\end{itemize}
In this work we restrict ourselves to non-spinng sources, where $\vec{S}_1=\vec{S}_2=0$.  Removing nuisance parameters requires either replacing them with their measured values or marginalizing over them.  The chirp mass is typically well measured by the detection pipeline, so we use the maximum likelihood value, $\hat{\mathcal{M}}$, for it.  We marginalize over the other variables using a uniform prior, i.e.
\begin{equation}
p(\alpha,\delta,D_L|d) = \int\, d\tau_\oplus p_U(\tau_\oplus) d\phi_c p_U(\phi_c) d\phi p_U(\psi)  d\cos(\iota) p_U(\cos(\iota)) p(\alpha,\delta,D_L,\tau_\oplus,\phi_c,\psi,\cos(\iota)|d),
\end{equation}
where $p_U(x)$ denotes a uniform distribution in $x$.

@@

\subsection{\acl{CRLB} and physical interpretation}

>>>>

The quantity of interest for any type of parameter estimation is the so-called {\it posterior} distribution, $p(\boldsymbol\theta|d)$, that describes the probability of obtaining the parameters, $\boldsymbol\theta$, given the data, $d$.  It is given by Bayes' theorem
\begin{equation}\label{bayes}
p(\boldsymbol\theta|d) = \frac{p(\boldsymbol\theta)p(d|\boldsymbol\theta)}{p(d)},
\end{equation}
where $p(\boldsymbol\theta)$ is the {\it prior} distribution describing previous knowledge of the paramters, $\boldsymbol\theta$, $p(d|\boldsymbol\theta)$ is the {\it likelihood} of obtaining the data, $d$, for the parameters, $\boldsymbol\theta$, and $p(d)$ is known as the {\it evidence}.  Procedures for computing the posterior typically rely on Markov chain Monte Carlo methods or nested sampling (something else?).  In this manuscript we present an alternative approach, taylored specifically to providing low-latency estimates of the sky location of and distance to the source.

In the full signal model the paramters, $\boldsymbol\theta$, are given by a 15-dimensional vector of quantities.  Of those, the quantities of interest are the right ascenion and declination, $(\alpha,\delta)$, and luminosity distance, $D_L$.  The remaining quantites, or {\it nuisance parameters} for the full signal model are:
\begin{itemize}
\item The chirp mass.  If $m_1$ and $m_2$ are the masses of the individual components then $\mathcal{M}=(m_1m_2)^{3/5}/(m_1+m_2)^{1/5}$.
\item The symmetric mass ratio, $\eta=(m_1m_2)/(m_1+m_2)^2$.
\item The 3-dimensional spin vectors of each components, $\mathbf{S}_1,\mathbf{S}_2$.
\item An arrival time, $\tau_\oplus$, taken here to be the time of arrival at the geocenter.
\item The inclination angle, $\iota$.
\item The polarization angle, $\psi$.
\item The coalescence phase, $\phi_c$.
\end{itemize}
In this work we restrict ourselves to non-spinng sources, where $\mathbf{S}_1=\mathbf{S}_2=0$.  Removing nuisance parameters requires either replacing them with their measured values or marginalizing over them.  The chirp mass is typically well measured by the detection pipeline, so we use the maximum likelihood value, $\hat{\mathcal{M}}$, for it.  We marginalize over the other variables using a uniform prior, i.e.
\begin{equation}
p(\alpha,\delta,D_L|d) = \int\, d\tau_\oplus p_U(\tau_\oplus) d\phi_c p_U(\phi_c) d\phi p_U(\psi)  d\cos(\iota) p_U(\cos(\iota)) p(\alpha,\delta,D_L,\tau_\oplus,\phi_c,\psi,\cos(\iota)|d),
\end{equation}
where $p_U(x)$ denotes a uniform distribution in $x$.

@@

\subsection{\ac{CRLB} and physical interpretation}
